Rhizaria in the oligotrophic ocean exhibit clear temporal and vertical variability.

### 2024-07-06 manuscript/04_discussion.qmd
change last sentence and add kim ref

The mesopelagic generally was home to known heterotrophic organisms, particularly for those which were constrained to exclusively occupy deeper waters. This is consistent with @blanco-bercial2022's observation of an auto-/mixotroph to heterotroph gradient in local protist community as well as global patterns in Rhizaria ecology \[@laget2024\]. The upper mesopelagic interestingly had relatively low total abundance. This low-abundance region likely reflects the dynamics of productivity and export throughout the water column. While productivity and thus sinking particles for flux feeders are high in the euphotic zone, much of this is attenuated throughout the epipelagic. So, while the base of the epipelagic may provide a rich feeding environment for Castanellidae, smaller protists, or heterotrophic bacteria (Figure 2F), the region from 200-500m might be otherwise food poor. Perhaps it is more advantageous for Rhizaria to situate deeper, in darker regions of the twilight zone where visual predators and vertically migrating organisms may not feed on them. Also it should be noted that Phaeodaria utilize silica to build their opaline tests, and silica concentrations started to increase around 500m (Figure 2G). However, $Si$ was not a significant term for any taxon-specific model in the mesopelagic. Aulosphaeridae was only found to have significant relationships, although weak fits, to particle concentration in the mesopelagic. In our study, while consistently observed, overall abundances of Aulosphaeridae were very low. In the Pacific Ocean, on California's Coast, much higher abundances of Aulosphaeridae have been reported \[@zasko2005; @biard2020\] and they have massive potential to impact silica export [@biard2018]. Coelodendridae were also seemingly restricted to the deeper section of the mesopelagic. This is interesting given that in the California Current, [@biard2020] found a bimodal distribution in Coelodendridae. There are several morphotypes corresponding to different taxa of Coelodendridae [@nakamura2015; @biard2020]. So it may be that only a few types of Coelodendridae were observed in this study, while the epipelagic variety was not. Alternatively, the lower epipelagic of the extremely productive California Current may provide adequate habitat for Coelodendridae, which is not available in the oligotrophic Sargasso Sea.

@@

@decelle2013 proposed a hypothetical life cycle for cyst-bearing (strictly heterotrophic) Acantharea. This hypothesized life cycle suggests that epipelagic Acantharea are adult populations, which form cysts that sink into the mesopelagic, then reproduce and rise. Furthermore, given that horizontal transfer of symbionts between generations of Acantharea is unlikely due to their spawning behavior, the newly spawned mesopelagic Acantharea are not necessarily required to rapidly return to the photic zone [@decelle2012; @decelle2013]. This hypothesis predicts that Acanthareas in the mesopelagic would be smaller [@decelle2013]. @marsbrisbin2020 provided some support for this hypothesis, with a significant decrease in Acantharea sizes with depth. Although the authors also observed low abundances in the mesopelagic and noted that the smaller sizes may be due to lower food availability [@marsbrisbin2020]. Since food is more scarce in the mesopelagic, nutritional quality lower, yet flux feeders would likely grow larger to increase their feeding range [@stukel2018; @biard2020]. In the data collecting in this study, Acanthareas in the mesopelagic were significantly smaller than the epipelagic, despite the other bimodal taxa (Foraminifera and Aulacanthidae) being significantly larger with depth. This provides added support for the hypothesis that cyst-forming Acantharea may utilize different sections of the water column throughout their life cycle. However to further investigate this, more work is needed with higher temporal and taxonomic resolution.

@@

The vertical partitioning documented in this study do support the hypothesis that mixotrophic rhizaria will occupy shallower waters while deeper waters are dominated by heterotrophy. However the degree to which mixotrophic Rhizaria in the euphotic zone rely on heterotrophy versus symbiosis is uncertain. Collodaria were recorded as consistent and dominant members of the near surface region. These organisms have the potential to contribute considerably to the otherwise low productivity of oligotrophic regions. However, their role in food webs is not well understood. While this study represents a step-forward in our understanding of Rhizaria ecology, continued research on Rhizaria is much needed to better understand their ecology. Particularly extended work describing interannual patterns in Rhizaria abundance. Also work defining biotic interactions, feeding rates, productivity, and life history are all rich fields of interest in Rhizaria.

>>>>

The mesopelagic generally was home to known heterotrophic organisms, particularly for those which were constrained to exclusively occupy deeper waters. This is consistent with @blanco-bercial2022's observation of an auto-/mixotroph to heterotroph gradient in local protist community as well as global patterns in Rhizaria ecology [@laget2024]. The upper mesopelagic interestingly had relatively low total abundance. This low-abundance region likely reflects the dynamics of productivity and export throughout the water column. While productivity and thus sinking particles for flux feeders are high in the euphotic zone, much of this is attenuated throughout the epipelagic. So, while the base of the epipelagic may provide a rich feeding environment for Castanellidae, smaller protists, or heterotrophic bacteria (Figure 2F), the region from 200-500m might be otherwise food poor. Perhaps it is more advantageous for Rhizaria to situate deeper, in darker regions of the twilight zone where visual predators and vertically migrating organisms may not feed on them. Also it should be noted that Phaeodaria utilize silica to build their opaline tests, and silica concentrations started to increase around 500m (Figure 2G). However, $Si$ was not a significant term for any taxon-specific model in the mesopelagic. Aulosphaeridae was only found to have significant relationships, although weak fits, to particle concentration in the mesopelagic. In our study, while consistently observed, overall abundances of Aulosphaeridae were very low. In the Pacific Ocean, on California's Coast, much higher abundances of Aulosphaeridae have been reported [@zasko2005; @biard2020] and they have massive potential to impact silica export [@biard2018]. Coelodendridae were also seemingly restricted to the deeper section of the mesopelagic. This is interesting given that in the California Current, [@biard2020] found a bimodal distribution in Coelodendridae. There are several morphotypes corresponding to different taxa of Coelodendridae [@nakamura2015; @biard2020]. So it may be that only a few types of Coelodendridae were observed in this study, while the epipelagic variety was not. Alternatively, the lower epipelagic of the extremely productive California Current may provide adequate habitat for Coelodendridae, which is not available in the oligotrophic Sargasso Sea.

@@

@decelle2013 proposed a hypothetical life cycle for cyst-bearing (strictly heterotrophic) Acantharea. This hypothesized life cycle suggests that epipelagic Acantharea are adult populations, which form cysts that sink into the mesopelagic, then reproduce and rise. Furthermore, given that horizontal transfer of symbionts between generations of Acantharea is unlikely due to their spawning behavior, the newly spawned mesopelagic Acantharea are not necessarily required to rapidly return to the photic zone [@decelle2012; @decelle2013]. This hypothesis predicts that Acantharea in the mesopelagic would be smaller [@decelle2013]. @marsbrisbin2020 provided some support for this hypothesis, with a significant decrease in Acantharea sizes with depth. Although the authors also observed low abundances in the mesopelagic and noted that the smaller sizes may be due to lower food availability [@marsbrisbin2020]. Since food is more scarce in the mesopelagic, nutritional quality lower [@kim2018], yet flux feeders would likely grow larger to increase their feeding range [@stukel2018; @biard2020]. In the data collecting in this study, Acantharea in the mesopelagic were significantly smaller than the epipelagic, despite the other bimodal taxa (Foraminifera and Aulacanthidae) being significantly larger with depth. This provides added support for the hypothesis that cyst-forming Acantharea may utilize different sections of the water column throughout their life cycle. However to further investigate this, more work is needed with higher temporal and taxonomic resolution.

@@

The vertical partitioning documented in this study do support the hypothesis that mixotrophic rhizaria will occupy shallower waters while deeper waters are dominated by heterotrophy. However the degree to which mixotrophic Rhizaria in the euphotic zone rely on heterotrophy versus symbiosis is uncertain. Collodaria were recorded as consistent and dominant members of the near surface region. These organisms have the potential to contribute considerably to the otherwise low productivity of oligotrophic regions. However, their role in food webs is not well understood. While this study represents a step forward in our understanding of Rhizaria, further research is needed to better characterize their ecology and role in biogeochemical cycling. To fully understand the functioning of the oceans, not only carbon or nitrogen, but other elements such as silica or strontium, extended work addressing the interannual patterns in Rhizaria, their biotic interactions, and the key biological rates that impact ocean biogeochemistry, will be required.
